---
title: "Cleaning and viz"
author: "Nerissa Nance"
date: "February 23, 2019"
output: html_document
---
#Data cleaning and visualiation

##Part 1: reduced dataset
The purpose of this file is to clean (as needed) as do some preliminary descriptive stats and visualizations on our variables of interest for the reduced dataset with 14 attributes.

Atribute information (from Kaggle[https://www.kaggle.com/ronitf/heart-disease-uci/version/1]):
 
> 1. age 
> 2. sex 
> 3. chest pain type (4 values) 
> 4. resting blood pressure 
> 5. serum cholestoral in mg/dl 
> 6. fasting blood sugar > 120 mg/dl
> 7. resting electrocardiographic results (values 0,1,2)
> 8. maximum heart rate achieved 
> 9. exercise induced angina 
> 10. oldpeak = ST depression induced by exercise relative to rest 
> 11. the slope of the peak exercise ST segment 
> 12. number of major vessels (0-3) colored by flourosopy 
> 13. thal: 3 = normal; 6 = fixed defect; 7 = reversable defect

```{r setup, include=FALSE}

library(rio)
library(tidyverse)
library(ggcorrplot)
library(umap)

data <- import("./data/heart.csv")

head(data)
names(data)
str(data)


#number of unique values in each column:
sapply(data, function(col){length(unique(col))})

#amount of missingness in each variable:
sapply(data, function(col){table(is.na(col))})
#none! weird, must have been imputed already at some point?


```

### Outcome and covariates 

```{r plots}

#look for outliers
lapply(data,function(col){
  summary(col)})


#outcome
ggplot(data, aes(x=as.factor(target),fill=as.factor(target)))+geom_bar()+labs(x="heart disease",title="Number of patients with and without outcome (heart disease)")+theme_classic()+ theme(legend.position = "none")


ggplot(data, aes(x=target, y=age, group=as.factor(target)))+geom_boxplot()
```

### Correlations and UMAP
```{r}
ggcorrplot(cor(data))


viz_umap <- umap(data)
layout <- as.data.frame(cbind(viz_umap$layout, viz_umap$data))

ggplot(data=layout,aes(V1,V2, color=as.factor(target))) + geom_point() +labs(x="UMAP1", y="UMAP2",title="Scatterplot of UMAP embeddings of heart disease data")+theme_classic()+scale_fill_continuous(name = "Presence of heart disease")
 

```


##Part 2: explore larger datasets

Here we will look at bringing back some of the variables that were initially in the larger heart disease datasets.

```{r, include=FALSE, eval=FALSE}

#load "updated" dataset
longbeach <- read_delim("http://archive.ics.uci.edu/ml/machine-learning-databases/heart-disease/long-beach-va.data", delim=" ")

hungarian <- read_delim("http://archive.ics.uci.edu/ml/machine-learning-databases/heart-disease/hungarian.data", delim=",")

cleveland <- read_delim("http://archive.ics.uci.edu/ml/machine-learning-databases/heart-disease/cleveland.data")


```
